---
title: "lab-03"
author: "Xingxin Li"
subtitle: "Lab-3"
date: "8/24/2020"
output: html_document
---
# Libraries
```{r message=FALSE, warning=FALSE, paged.print=FALSE}
# SPDS
library(tidyverse)
library(sf)
library(units)
# Data
library(USAboundaries)
library(rnaturalearthdata)
library(rnaturalearth)
# Visualization
library(gghighlight)
library(ggrepel)
library(knitr)

source = "/Users/xingxin/Github/geog176a-summer-2020-lab1/"

```

# Background
In this lab, 4 main skills are covered:

 1. __Ingesting/Building__ _sf_ objects from R packages and CSVs. (Q1)
 2. __Manipulating__ geometries and coordinate systems (Q2)
 3. __Calculaing__ distances (Q2)
 4. __Building__ maps using ggplot (Q3)
 
# Question 1
### _1.1 Define a Projection_
```{r}
eqdc = '+proj=eqdc +lat_0=40 +lon_0=-96 +lat_1=20 +lat_2=60 +x_0=0 +y_0=0 +datum=NAD83 +units=m +no_defs'
```
Description of  this projection based on the parameters:

* projection name: eqdc
* Latitude of origin: 40
* Longitude of origin: -96
* Latitude of first standard parallel: 20
* Latitude of second standard parallel: 60
* False Easting: 0
* False Northing: 0
* Datum: NAD83
* Units: m
 
### _1.2 - Get USA state boundaries_
```{r}
state = USAboundaries::us_states(resolution = "low") %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii")) %>%
  st_transform(eqdc)
```

### _1.3 - Get country boundaries for Mexico, the United States of America, and Canada_
```{r}
country = rnaturalearth::countries110
country_sf = st_as_sf(country) %>%
  filter(admin %in% c("United States of America", "Mexico", "Canada")) %>%
  st_transform(eqdc)
```

### _1.4 - Get city locations from the CSV file_
```{r}
city = read_csv(paste(source,"uscities.csv",sep = "")) %>%
  select(city, state_name, county_name, population, lat, lng)
city_sf = st_as_sf(city, coords = c("lng", "lat"), crs = 4326) %>%
  filter(!state_name %in% c("Puerto Rico", "Alaska", "Hawaii")) 
city_eqdc = st_transform(city_sf, eqdc) 
   
```

# Question 2
### _2.1 - Distance to USA Border (coastline or national) (km)_
``` {r}
# Union States: state boundaries are resolved
usa_geom = state$geometry
usa_u_ml = st_union(usa_geom)  %>%
    st_cast("MULTILINESTRING")
# Combine States: state boundaries are preserved
usa_c_ml = st_combine(usa_geom)  %>%
    st_cast("MULTILINESTRING")

# Distance to US Border
city_eqdc = city_eqdc %>%
  mutate(dist_border = st_distance(city_eqdc, usa_u_ml)) %>%
  mutate(dist_border = units::set_units(dist_border, "km"))

# Five Cities to US Border 
city_eqdc %>% select(city,state_name, dist_border) %>% slice_max(dist_border, n = 5) %>% 
  st_drop_geometry() %>% 
  knitr::kable(caption = "Five Cities Farthest from US Border",
             col.names = c("City", "State", "Distance to US Border")) 

```

### _2.2 - Distance to States (km)_
``` {r}
# The distance of each city to the nearest state boundary
city_eqdc = city_eqdc %>%
  mutate(dist_state = st_distance(city_eqdc, usa_c_ml)) %>%
  mutate(dist_state = units::set_units(dist_state, "km"))

# Five Cities to US State
city_eqdc %>% select(city,state_name, dist_state) %>% slice_max(dist_state, n = 5) %>% 
  st_drop_geometry() %>% 
knitr::kable(caption = " Five Cities Farthest from State Border",
             col.names = c("City", "State", "Distance to State Border"))

```

### _2.3 - Distance to Mexico (km)_
``` {r}
# The distance of each city to the Mexican border
city_eqdc = city_eqdc %>%
  mutate(dist_Mexico = st_distance(city_eqdc, filter(country_sf, admin == "Mexico"))) %>%
  mutate(dist_Mexico = units::set_units(dist_Mexico, "km"))

# Five Cities to Mexican
city_eqdc %>% select(city,state_name, dist_Mexico) %>% slice_max(dist_Mexico, n = 5) %>% 
  st_drop_geometry() %>% 
knitr::kable(caption = " Five Cities Farthest from Mexican Border",
             col.names = c("City", "State", "Distance to Mexican Border"))

```

### _2.4 - Distance to Canada (km)_
``` {r}
# The distance of each city to the Canadian border
city_eqdc = city_eqdc %>%
  mutate(dist_Canada = st_distance(city_eqdc, filter(country_sf, admin == "Canada"))) %>%
  mutate(dist_Canada = units::set_units(dist_Canada, "km"))

# Five Cities to Canada
city_eqdc %>% select(city,state_name, dist_Canada) %>% slice_max(dist_Canada, n = 5) %>% 
  st_drop_geometry() %>% 
knitr::kable(caption = " Five Cities Farthest from Canadian Border",
             col.names = c("City", "State", "Distance to Canadian Border"))
```



# Question 3
### _3.1 Data_
``` {r}
# 10 largest cities
largest_cities10 = city_eqdc %>%
  slice_max(population, n = 10)

ggplot() +
  geom_sf(data = country_sf) +
  geom_sf(data = state, lty = 3, size = .5) +
  geom_sf(data = largest_cities10, size = 1.5,color = "red") +
  ggrepel::geom_label_repel(
    data = largest_cities10,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 3) +
  labs(title = "10 Largest USA Cities (by population)",
       x = " ",
       y = " ") +
  ggthemes::theme_map() 

```

### _3.2 City Distance from the Border_
``` {r}
# 5 farthest cities
farthest_cities5 = city_eqdc %>%
  slice_max(dist_border, n = 5)

ggplot() +
  geom_sf(data = city_eqdc, aes(col= as.numeric(dist_border))) +
  scale_color_gradient(low = 'blue', high = "red", name = "Distance (km)") +
  geom_sf(data = state, size = 0.5,lty = 3, fill = "NA",) +
  geom_sf(data = farthest_cities5, size = 1, color = "black") +
  ggrepel::geom_label_repel(
    data = farthest_cities5,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 3) +
    labs(title = "City Distance from National Border and The Five Farthest Cities",
       x = " ",
       y = " ") + 
  ggthemes::theme_map() 
```

### _3.3 City Distance from Nearest State_
``` {r}
farthest_cities5_state = city_eqdc %>%
  slice_max(dist_state, n = 5)

ggplot() +
  geom_sf(data = city_eqdc, aes(col= as.numeric(dist_state))) +
  scale_color_gradient(low = 'blue', high = "red", name = "Distance (km)") +
  geom_sf(data = state, size = 0.5,lty = 3, fill = "NA",) +
  geom_sf(data = farthest_cities5_state, size = 1, color = "black") +
  ggrepel::geom_label_repel(
    data = farthest_cities5_state,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 3) +
    labs(title = "City Distance from Nearest State Border and The Five Farthest Cities",
       x = " ",
       y = " ") + 
  ggthemes::theme_map() 


```

### _3.4 Equidistance boundary from Mexico and Canada_
``` {r}



city_eqdc = city_eqdc %>%
  mutate(diff = abs(dist_Canada - dist_Mexico)) 
  
equal_cities3 = city_eqdc %>%
  filter(as.numeric(diff) < 100 ) %>%
  slice_max(population, n = 5)

ggplot() +
  geom_sf(data = city_eqdc, aes(col= as.numeric(diff))) +
  scale_color_gradient(low = 'grey', high = "red", name = "Distance (km)") +
  gghighlight(as.numeric(diff)<100) +
  geom_sf(data = state, size = 0.5,lty = 3, fill = "NA",) +
  geom_sf(data = equal_cities3, size = 1, color = "black") +
  ggrepel::geom_label_repel(
    data = equal_cities3,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 3) +
    labs(title = " Cities that are equal distance from the Canadian AND Mexican border ± 100 km",
         subtitle = "5 Most Populous Cities in This Zone",
       x = " ",
       y = " ") + 
  ggthemes::theme_map() 

```


# Question 4
## Real World Application
### _4.1 Quantifing Border Zone_
``` {r}
# How many cities are in this 100 mile zone? (100 miles ~ 160 kilometers)

# How many people live in a city within 100 miles of the border?

# What percentage of the total population is in this zone?
zone = city_eqdc %>%
  mutate(total_pop = sum(population)) %>%
  filter(as.numeric(dist_border) <= 160) %>%
  summarise(number_cities = n(), sum_pop = sum(population), percent = 100 * (sum_pop/total_pop[1])) %>%
  st_drop_geometry() 

knitr::kable(zone,
             caption = "Cities Within 100 Miles of State Border",
             col.names = c("Number of Cities", "Number of People", "Percent of Population(%)"))

```
### Does it match the ACLU estimate in the link above?
The ACLU states that "Roughly two-thirds of the United States' population lives within the 100-mile zone—that is, within 100 miles of a U.S. land or coastal border. That's about 200 million people."

(1)The total number of people live in a city within 100 miles of the border calculated above is:

* `r zone$sum_pop` 

It's about 260 million, which is higher than the ACLU's estimation.

(2)The percent Percent of Population  calculated above is:

* `r zone$percent` 

which is close to the ACLU's estimation.


### _4.2 Mapping Border Zone_
``` {r}

pop_zone10 = city_eqdc %>%
  filter(as.numeric(dist_border) <= 160) %>%
  select(city, population) %>%
  slice_max(population, n = 10)

ggplot() +
  geom_sf(data = city_eqdc, aes(col= as.numeric(dist_border))) +
  scale_color_gradient(low = 'orange', high = "darkred", name = "Distance (km)") +
  gghighlight(as.numeric(city_eqdc$dist_border) <= 160) +
  geom_sf(data = state, size = 0.5,lty = 3, fill = "NA",) +
  geom_sf(data = pop_zone10, size = 1, color = "black") +
  ggrepel::geom_label_repel(
    data = pop_zone10,
    aes(label = city, geometry = geometry),
    stat = "sf_coordinates",
    size = 3) +
    labs(title = " Cities within the 100 Mile Zone",
         subtitle = "10 Most Populous Cities in This Zone",
       x = " ",
       y = " ") + 
  ggthemes::theme_map() 
  
```
