---
title: "lab5"
author: "Xingxin Li"
date: "9/4/2020"
output: html_document
---
### _Libraries_
``````{r, message = FALSE, warning = FALSE}
library(raster) 
library(tidyverse)  
library(getlandsat)  
library(sf)  
library(mapview)
library(osmdata)
library(getlandsat)
library(leaflet)
library(units)
library(factoextra)
```


# _Question 1 : Find AOI_

``````{r, message = FALSE, warning = FALSE}

bb = read.csv("/Users/xingxin/Github/geog176a-summer-2020-lab1/uscities.csv") %>%
  filter(city == "Palo") %>%
  st_as_sf(coords = c("lng","lat"), crs = 4326) %>%
  st_transform(5070) %>%
  st_buffer(5000) %>%
  st_bbox() %>%
  st_as_sfc() %>%
  st_as_sf()

mapview(bb)
```

# _Question 2_
### _Q2.1 landset discovery space and time_
```{r, message = FALSE, warning = FALSE}

bbwgs = bb %>% st_transform(4326)

bb = st_bbox(bbwgs)

scenes = lsat_scenes()

down = scenes %>%
  filter(min_lat <= bb$ymin, max_lat >= bb$ymax,
         min_lon <= bb$xmin, max_lon >= bb$xmax,
         as.Date(acquisitionDate) == as.Date("2016-09-26"))

write.csv(down, file = "/Users/xingxin/Github/geog176a-summer-2020-lab1/palo-flood-scene.csv")
```


### _Q2.2-2.3 crop_
```{r, message = FALSE, warning = FALSE}
meta = read_csv("/Users/xingxin/Github/geog176a-summer-2020-lab1/palo-flood-scene.csv")

files = lsat_scene_files(meta$download_url) %>%
  filter(grepl(paste0("B", 1:6, ".TIF$", collapse = "|"), file)) %>%
  arrange(file) %>%
  pull(file)

st = sapply(files, lsat_image)

s = stack(st) %>% 
  setNames(c(paste0("band", 1:6)))

cropper = bbwgs %>% 
  st_transform(crs(s)) 

r = crop(s, cropper)

```

### _Q2.3 Question_


The dimension I stacks has 7811 rows and 7681 columns, 6 layers. The CRS is WGS84, the cell resolution is x =30, y = 30. 


### _Q2.4 Question_

The dimension I cropped has 340 rows and 346 columns, 6 layers. The CRS is WGS84, the cell resolution is x =30, y = 30. 

# _Question 3_
### _Q3 plotting_
```{r, message = FALSE, warning = FALSE}
#rgb
par(mfrow = c(1,2))
plotRGB(r, r = 4, g = 3, b = 2, stretch = "lin")
plotRGB(r, r = 4, g = 3, b = 2, stretch = "hist")

#nir-rg
par(mfrow = c(1,2))
plotRGB(r, r = 5, g = 4, b = 3, stretch = "lin")
plotRGB(r, r = 5, g = 4, b = 3, stretch = "hist")

#nir-swir1-r
par(mfrow = c(1,2))
plotRGB(r, r = 6, g = 5, b = 4, stretch = "lin")
plotRGB(r, r = 6, g = 5, b = 4, stretch = "hist")

#g-nir-b,my choice
par(mfrow = c(1,2))
plotRGB(r, r = 3, g = 6, b = 2, stretch = "lin")
plotRGB(r, r = 3, g = 6, b = 2, stretch = "hist")

```
# _Question 4_
### _Q4.1 Local Operation_
```{r, message = FALSE, warning = FALSE}
#ndvi
ndvi = (r$band5 - r$band4) / (r$band5 + r$band4)
palette1 = colorRampPalette(c("blue", "white", "red"))
plot(ndvi, col = palette1(256))

#ndwi
ndwi = (r$band3 - r$band5) / (r$band3 + r$band5)
palette2 = colorRampPalette(c("blue", "white", "red"))
plot(ndwi, col = palette2(256))

#mndvi
mndvi = (r$band3 - r$band6) / (r$band3 + r$band6)
palette3 = colorRampPalette(c("blue", "white", "red"))
plot(mndvi, col = palette3(256))

#wri
wri = (r$band3 + r$band4) / (r$band5 + r$band4)
palette4 = colorRampPalette(c("blue", "white", "red"))
plot(wri, col = palette4(256))

#swi
swi = 1 / sqrt(r$band2 - r$band6)
palette5 = colorRampPalette(c("blue", "white", "red"))
plot(swi, col = palette5(256))
```
### _Q4.2 Thresholding_
```{r, message = FALSE, warning = FALSE}
#ndvi
thresholdingN = function(x){ifelse(x <= 0, 1, NA)}

flood = calc(ndvi, thresholdingN)
plot(flood, col = "blue")

#ndwi
thresholdingD = function(x){ifelse(x >= 0, 1, NA)}

flood2 = calc(ndwi, thresholdingD)
plot(flood2, col = "blue")

#mndvi
thresholdingM = function(x){ifelse(x >= 0, 1, NA)}

flood3 = calc(mndvi, thresholdingM)
plot(flood3, col = "blue")

#wri
thresholdingW = function(x){ifelse(x >= 1, 1, NA)}

flood4 = calc(wri, thresholdingW)
plot(flood4, col = "blue")

#swi
thresholdingS = function(x){ifelse(x <= 5, 1, NA)}

flood5 = calc(swi, thresholdingS)
plot(flood5, col = "blue")


#stack
flood_stacks = stack(flood, flood2, flood3, flood4, flood5)%>%
  setNames(c("ndvi", "ndwi", "mndvi", "wri", "swi" ))
plot(flood_stacks, col ="blue")
```

### _4.4 Question_

In these five images, they all showing the floods around the river, However, the ndvi, ndwi, and wri look familiar compared with mndvi and swi because of the calculation and water threshold are different.


# _Question 5_
### _Q5.1 set.seed_
```{r, message = FALSE, warning = FALSE}
set.seed(09032020) 
v = getValues(r) 

idx = which(!is.na(v))
v = na.omit(v)
```

### _5.2 extract and check_
```{r, message = FALSE, warning = FALSE}
k2 = kmeans(v, 12, iter.max = 100)

knew_raster = ndvi
values(knew_raster) = NA
knew_raster[idx] <- k2$cluster

plot(knew_raster, col = RColorBrewer::brewer.pal(5, "Spectral"))
```

### _5.2 Question_

The data was extract in 2 datasets and one of the datasets has 340 346 6 in the sets and the which is tolds there's 117640 total rows and column in 6 layers.

### _5.3 flood category_
```{r, message = FALSE, warning = FALSE}
k3 = kmeans(v, 12, iter.max = 100)
knew_raster3 = knew_raster
values(knew_raster3) = NA
knew_raster3[idx] <- k3$cluster


flood_raster_stacks = addLayer(flood_stacks, knew_raster3)
plot(flood_raster_stacks)
```


# _Question 6_
### _6 Plot-Print_
```{r, message = FALSE, warning = FALSE}
andvi = cellStats(flood, stat = 'sum')*900
andwi = cellStats(flood2, stat = 'sum')*900
amndvi = cellStats(flood3, stat ='sum')*900 
awri = cellStats(flood4, stat = 'sum')*900 
aswi = cellStats(flood5, stat = 'sum')*900
aknew3 = cellStats(knew_raster3, stat = 'sum')*900

flood_summary = bind_rows(andvi = cellStats(flood, stat = 'sum'), 
andwi = cellStats(flood2, stat = 'sum'),
amndvi = cellStats(flood3, stat ='sum'), 
awri = cellStats(flood4, stat = 'sum'), 
aswi = cellStats(flood5, stat = 'sum'),
aknew3 = cellStats(knew_raster3, stat = 'sum'))

knitr::kable(flood_summary,
caption = "Flood Summary",
format.args = list(big.mark = ",")) %>%
kableExtra::kable_styling("basic", full_width = TRUE, font_size = 16)
mapview(flood_raster_stacks, col = RColorBrewer::brewer.pal(5, "Spectral"))

```
### _Question_


The reason I thinks the cell values not an even number is because the calc function has determine the visualization as using the integer number to making the resolution filling in the map and the map fill out the area with a similarity color to make the visualization more efficient.


